---
title: "SRTR Data Preparation Template"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
# Load essential libraries
library(tidyverse)
library(haven)  # For reading SRTR SAS files
library(lubridate)  # For date handling

  # Specify file paths
cand_kipa_path <- "path/to/CAND_KIPA.sas7bdat"
tx_hr_path <- "path/to/TX_HR.sas7bdat"

# Load datasets
cand_kipa <- read_sas(cand_kipa_path)
tx_hr <- read_sas(tx_hr_path)

# FIlter to cohort
cand_thor <- cand_thor |> 
  subset(CAN_LISTING_DT >= as.Date("2019-01-01") & CAN_LISTING_DT <= as.Date("2022-12-31") & WL_ORG == "HR")

# IF MULTIPLE LISTING IS AN ISSUE
```{r}
cand_thor <- cand_thor %>% ##cand_thor has same cohort, adds waitlist_end_date column
  mutate(waitlist_end_date = case_when(
    is.na(REC_TX_DT) == FALSE ~ REC_TX_DT, ##look at transplant date first
    is.na(CAN_REM_DT) == FALSE ~ CAN_REM_DT, ##then look at removal date
    is.na(CAN_LAST_INACT_STAT_DT) == FALSE & CAN_LAST_INACT_STAT_DT > CAN_LAST_ACT_STAT_DT ~ CAN_LAST_INACT_STAT_DT,##then inactive date if after last active date
    !is.na(CAN_LAST_ACT_STAT_DT) ~ CAN_LAST_ACT_STAT_DT, ##then last active
    is.na(CAN_LAST_ACT_STAT_DT) & !is.na(CAN_LAST_INACT_STAT_DT) ~ CAN_LAST_INACT_STAT_DT,
    TRUE ~ CAN_LAST_ACT_STAT_DT)) ##if all else fails, use last active date

```

##single registrations - new columns num_list, min_list_date, wait_time, outcome

```{r}
single_registrations <- cand_thor %>%
  group_by(PERS_ID) %>% ##grouping by PERS_ID - ?
  mutate(num_list = n()) %>%
  filter(num_list == 1) %>%
  ungroup() %>%
  mutate(min_list_date = CAN_LISTING_DT,  
         wait_time = waitlist_end_date - min_list_date, ##using end dates established in last code chunk
         outcome = case_when(
           DON_TY == "C" ~ "transplanted",
           !is.na(CAN_REM_CD) ~ "removed/died",  ##do we want all other CAN_REM_CD codes here? or just 8 and 13 (like used to define the outcome below)
           TRUE ~ "censored"
         ))

```

##multiple registrations - new columns num_list, list_type (concurrent or sequential), num_tx; fills missing REC_TX_DT

```{r}
multiple_registrations <- cand_thor %>%
  filter(!PX_ID %in% single_registrations$PX_ID) %>%
  group_by(PERS_ID) %>%
  arrange(PERS_ID, CAN_LISTING_DT) %>%
  mutate(num_list = n())

candidates_w_multiple_registrations <- multiple_registrations %>% pull(PERS_ID) %>% unique() %>% length() ##number of candidates with multiple registrations

avg_num_registrations_multiple <- mean(multiple_registrations %>% group_by(PERS_ID) %>% filter(row_number() == 1) %>% pull(num_list)) ##for candidates with multiple registrations, the average number of registrations

multiple_registrations <- multiple_registrations %>%
  mutate(list_type = case_when(
    CAN_LISTING_DT < lag(waitlist_end_date) ~ "concurrent",
    waitlist_end_date > lead(CAN_LISTING_DT) ~ "concurrent",
    TRUE ~ "sequential")) %>%
  mutate(REC_TX_DT = as.Date(REC_TX_DT)) %>%
  mutate(num_tx = length(unique(na.omit(REC_TX_DT)))) %>%
  fill(REC_TX_DT, .direction='downup') ##this line is causing issues with negative dial_time values
```

```{r}
multiple_registrations <- multiple_registrations[order(multiple_registrations$PERS_ID, multiple_registrations$waitlist_end_date), ] ##orders by PERS_ID then waitlist_end_date

##multiple registrations - number sequential transplants, correct concurrent listing transplant dates

multiple_registrations$transplant_num <- 1 ##new column transplant_num, setting all at 1 for now
for (i in 2:nrow(multiple_registrations)) {
  if (!is.na(multiple_registrations$PERS_ID[i-1]) && 
      !is.na(multiple_registrations$PERS_ID[i]) && 
      !is.na(multiple_registrations$REC_TX_DT[i-1]) && 
      !is.na(multiple_registrations$REC_TX_DT[i]) &&
      multiple_registrations$PERS_ID[i-1] == multiple_registrations$PERS_ID[i] &&
      multiple_registrations$REC_TX_DT[i-1] != multiple_registrations$REC_TX_DT[i]) {
    
    multiple_registrations$transplant_num[i] = multiple_registrations$transplant_num[i-1] + 1
  }
}

for (i in 2:nrow(multiple_registrations)) {
  if (!is.na(multiple_registrations$PERS_ID[i-1]) && 
      !is.na(multiple_registrations$PERS_ID[i]) && 
      multiple_registrations$PERS_ID[i-1] == multiple_registrations$PERS_ID[i] &&
      !is.na(multiple_registrations$transplant_num[i-1]) && 
      multiple_registrations$transplant_num[i-1] != multiple_registrations$transplant_num[i] &&
      multiple_registrations$transplant_num[i-1] != 1) {
    
    multiple_registrations$transplant_num[i] = multiple_registrations$transplant_num[i-1]
  }
}

multiple_registrations$transplant_num[multiple_registrations$list_type == 'sequential'] <- 0 ##when list_type is sequential, transplant_num is set to 0. I still don't understand this step either.

for(i in 1:(nrow(multiple_registrations)-1)) { ##loop from second to lat row
  if(multiple_registrations$PERS_ID[i] == multiple_registrations$PERS_ID[i+1] &
     multiple_registrations$list_type[i] == 'concurrent' & multiple_registrations$list_type[i+1] == 'concurrent' &
     !is.na(multiple_registrations$REC_TX_DT[i]) & !is.na(multiple_registrations$REC_TX_DT[i+1]) &
     multiple_registrations$REC_TX_DT[i] < multiple_registrations$REC_TX_DT[i+1] ) {
    
    multiple_registrations$REC_TX_DT[i] <- multiple_registrations$REC_TX_DT[i+1] ##if PERS_ID is the same as the NEXT row, both have concurrent listings, transplant dates for both are not NA, and transplant date is earlier than the next row's transplant date - update to next row's transplant date
  }}

sequential_lists <- multiple_registrations %>%
  filter(list_type == "sequential") %>%
  mutate(min_list_date = CAN_LISTING_DT,
         wait_time = waitlist_end_date - min_list_date,
         outcome = case_when(
           DON_TY == "C" ~ "transplanted",
           is.na(CAN_REM_CD) == FALSE ~ "removed/died", ##again, should this be all CAN_REM_CD? or just 8 and 13 (since we included the rest in censored)
           TRUE ~ "censored"
         ))  ##same thing as done above for single registrations

max_retransplants <- max(multiple_registrations$transplant_num) ##highest number of transplants for a single person

multiple_registrations <- multiple_registrations %>% ##for candidates with multiple registrations, adds columns min_list_date, wait_time
  group_by(PERS_ID, transplant_num) %>%
  mutate(min_list_date = min(CAN_LISTING_DT, na.rm=T),
         wait_time = waitlist_end_date - min_list_date)

multiple_registrations <- multiple_registrations %>% mutate

collapsed_concurrent_registrations <- NULL ##establishes object to store coming data
for(i in 1:max_retransplants) { ##loop from first row to max_transplants
  
  collapsed_concurrent_registrations <- rbind(collapsed_concurrent_registrations, 
        
  multiple_registrations %>%
    filter(list_type == "concurrent" & transplant_num == i) %>% ##filters for concurrent listings where translpant_num is equal to current value of i (?)
    mutate(DON_TY = ifelse(DON_TY == "", NA, DON_TY),
           last_wait_date = max(waitlist_end_date, na.rm = TRUE)) %>% ##if DON_TY is empty, fill with NA, set last_wait_date to waitlist_end_date (why?)
    fill(REC_TX_DT, .direction = "up") %>%
    fill(DON_TY, .direction = "up") %>%
    fill(DONOR_ID, .direction = "up") %>%
    fill(CAN_REM_CD, .direction = "up") %>% ##fills missing values upwards
    mutate(wait_time = case_when(
      is.na(REC_TX_DT) == FALSE & transplant_num != '0' ~ REC_TX_DT- min_list_date, ### Ignore non-transplanted rows
      TRUE ~ last_wait_date - min_list_date),
      outcome = case_when(
        DON_TY == "C" ~ "DDKT",
        DON_TY == "L" ~ "LDKT",
        is.na(CAN_REM_CD) == FALSE ~ "removed/died", ##same question - CAN_REM_CD should be all? or just 8, 13?
        TRUE ~ "censored")   ##does the same we already did for single registrations and sequential listings
    ) %>%
    select(-c(waitlist_end_date, CAN_LISTING_DT, CAN_REM_DT)) %>% ##drops columns waitlist_end_date, CAN_LISTING_DT, CAN_REM_DT (just for cleanliness?)
    filter(row_number() ==1) %>% ##for concurrent listings, keeps only first row 
    
    mutate(last_wait_date = case_when(
      REC_TX_DT < last_wait_date ~ REC_TX_DT,
      TRUE ~last_wait_date)))}
```

##recreate CAN_LISTING_DT - to fix issue of NA values later on

```{r}
collapsed_concurrent_registrations <- collapsed_concurrent_registrations %>% mutate(CAN_LISTING_DT = min_list_date)
```
##recombine separated data frames

```{r}
cand_thor <- bind_rows(single_registrations %>% ungroup(), 
                          sequential_lists %>% ungroup(), 
                          collapsed_concurrent_registrations %>% ungroup())
```

# Define relevant date and time variables

cand_thor <- cand_thor |> 
  mutate(
    start_date = as.Date(CAN_LISTING_DT),
    transplant_date = as.Date(REC_TX_DT),
    removal_date = as.Date(CAN_REM_DT),
    last_active_date = as.Date(CAN_LAST_ACT_STAT_DT),
    last_inactive_date = as.Date(CAN_LAST_INACT_STAT_DT),
    death_date = as.Date(CAN_DEATH_DT),
    death_date_max = pmax(CAN_DEATH_DT, PERS_SSA_DEATH_DT, PERS_OPTN_DEATH_DT, na.rm=T)) %>%
  
  mutate(transplant_time = ifelse(!is.na(transplant_date), 
                                  as.numeric(transplant_date - start_date, units='days'), NA)) %>%
  mutate(death_time = ifelse(!is.na(death_date_max), as.numeric(death_date_max - start_date, units='days'), NA)) %>%
  mutate(removal_time = as.numeric(pmax(removal_date, last_active_date, last_inactive_date, na.rm = T) - 
                                     start_date, units='days'))

# Define survival time, censoring at a certain length of time
Survival time defined as time to transplant, death, or removal from waitlist, whichever comes first. Units = days (from previous code)

```{r survival time}
 cand_thor <- 
  cand_thor |> 
  mutate(
    survival_time = pmin(transplant_time, death_time, removal_time, na.rm = TRUE),
    survival_time = case_when(
      survival_time > 1095 ~ 1095,
      .default = survival_time
    ))
```
# Define statuses (will change project to project)
```{r censoring}
cand_thor <- cand_thor |> 
  group_by(PX_ID) |> 
    mutate(status = case_when(
      survival_time == 1095 ~ 0, 
      survival_time == death_time ~ 2,
      survival_time == transplant_time  ~ 1,
      !is.na(CAN_REM_CD) & !(CAN_REM_CD %in% c(4,8)) ~ 3,
      .default = 0
      ))

cand_thor$status <- factor(cand_thor$status,
                           levels = c(0,1,2,3),
                           labels = c("censor",  "transplant","death", "removal"))


```

# raw EPTS formula

```{r}
df_intervals <- df_intervals %>% 
  mutate(raw_epts = 
           0.047*pmax(age - 25, 0) - 
           0.015*(diabetes==1)*pmax(age - 25, 0) +
           0.398*(previous_TX==1) - 0.237*(diabetes==1)*(previous_TX==1) +
           0.315*log(dialysis_time + 1) - 0.099*(diabetes==1)*log(dialysis_time + 1) +
           0.130*(dialysis_time == 0) - 0.348*(diabetes==1)*(dialysis_time == 0) +  
           1.262*(diabetes==1))
```
# KDRI formula - CHANGE IF RACE AND/OR HCV ARE REMOVED FROM SCORE
data <- data %>%
  mutate(
    kdri_rao = exp(
      0.0128 * (age - 40) +
      -0.0194 * (age - 18) * (age < 18) +
      0.0107 * (age - 50) * (age > 50) +
      -0.0464 * (height - 170) / 10 +
      -0.0199 * (weight - 80) / 5 * (weight < 80) +
      0.1790 * african_american +
      0.1260 * hypertension +
      0.1300 * diabetes +
      0.0881 * cva +
      0.2200 * (creatinine - 1) +
      -0.2090 * (creatinine - 1.5) * (creatinine > 1.5) +
      0.2400 * hcv +
      0.1330 * dcd
    )
  )

# Define the median KDRI_RAO for scaling (reference value)
median_kdri_rao <- median(kdri_rao_prev_year)  # Replace with actual reference value if available

# Calculate scaled KDRI
data <- data %>%
  mutate(kdri = kdri_rao / median_kdri_rao)

# Calculate epts/kdpi using mapping tables
```{r}
epts_mapping <- read_csv("epts_mapping_table.csv", col_names = c("min_raw", "max_raw", "epts_value"))
kdpi_mapping <-read_csv("epts_mapping_table.csv", col_names = c("min_kdri", "max_kdri", "kdpi_value"))

# Function to map raw EPTS to EPTS value
calculate_epts <- function(raw_epts, mapping_table) {
  result <- sapply(raw_epts, function(x) {
    match_row <- mapping_table %>%
      filter(x >= min_raw & x < max_raw)
    if (nrow(match_row) > 0) {
      return(match_row$epts_value)
    } else {
      return(NA)  # Return NA if no match is found
    }
  })
  return(result)
}

calculate_kdpi <- function(kdri, mapping_table) {
  result <- sapply(kdri, function(x) {
    match_row <- mapping_table %>%
      filter(x >= min_kdri & x < max_kdri)
    if (nrow(match_row) > 0) {
      return(match_row$kdpi_value)
    } else {
      return(NA)  # Return NA if no match is found
    }
  })
  return(result)
}

data <- data %>%
  mutate(epts = calculate_epts(raw_epts, epts_mapping),
        kdpi = calculate_kdpi(kdri, kdpi_mapping))
```

# IF LISTINGS FOR OTHER ORGANS ARE RELEVANT
```{r}
check_overlap <- function(df, other_df) {
  # Filter for common PERS_ID
  common_ids <- intersect(df$PERS_ID, other_df$PERS_ID)
  
  # Subset datasets to only include common PERS_ID
  filtered_df <- df %>% filter(PERS_ID %in% common_ids)
  filtered_other_df <- other_df %>% filter(PERS_ID %in% common_ids)
  
  # Join heart_df with the filtered other_df for date comparison
  result <- filtered_df %>%
    left_join(filtered_other_df, by = "PERS_ID", suffix = c("_main", "_other")) %>%
    filter(
      # Case 1: Overlap with removal date
      (CAN_LISTING_DT_main >= CAN_LISTING_DT_other & 
       CAN_LISTING_DT_main <= CAN_REM_DT_other) |
      # Case 2: No removal date, listing dates must match or overlap
      (is.na(CAN_REM_DT_other) & CAN_LISTING_DT_main >= CAN_LISTING_DT_other)
    )
  
  return(result)
}

overlap <- check_overlap(data, other_organ_data)
overlap <- overlap %>% select(PERS_ID, WL_ORG_main, WL_ORG_other)
overlap <- as.data.frame(overlap %>% group_by(PERS_ID) %>% distinct())

# One row per PERS_ID (combines multiple other organ listings into one)
consolidated_overlap <- overlap_results %>%
  group_by(PERS_ID) %>%
  dplyr::summarize(WL_ORG_combined = paste(unique(WL_ORG_other), collapse = ", "), .groups = "drop")
consolidated_overlap


# Save data for import in analysis/figures files

