---
title: "National US-CRS Allocation System"
subtitle: "Simplified distance-free US-CRS based heart allocation"
output: html_notebook
---

This notebook implements a simplified national US-CRS based heart allocation system that ignores geographic distance while maintaining essential prioritization for pediatric status, donor age groups, and ABO compatibility.

## Key Features:
- **No Distance Constraints**: National allocation pool for all organs
- **Simplified Classifications**: 24 main classes (vs 172 in full system)
- **Pediatric Protection**: Pediatric candidates prioritized for pediatric donors
- **US-CRS Priority**: Higher scores indicate higher transplant need
- **Blood Type Compatibility**: Primary matches preferred over secondary

## Output:
- `national_uscrs_classification`: Classification number (1-24+) per national allocation
- `national_uscrs_sequence`: New sequence number based on national US-CRS rules

```{r setup}
# Load essential libraries
library(tidyverse)
library(haven)      # For SAS file reading
library(lubridate)  # For date handling
library(here)       # For file path management
```

# Load match run data with US-CRS scores
```{r load-data}
# Load the processed match run data with US-CRS scores and waiting times
load(here("data", "heart_only_match_run_with_uscrs.Rdata"))

print(paste("Loaded", nrow(heart_only_match_run_with_uscrs), "match run records"))

# Verify waiting time column is available
if("waiting_time_days" %in% colnames(heart_only_match_run_with_uscrs)) {
  waiting_summary <- heart_only_match_run_with_uscrs %>%
    filter(!is.na(waiting_time_days) & waiting_time_days >= 0) %>%
    summarise(
      records_with_waiting_time = n(),
      mean_waiting = round(mean(waiting_time_days), 1),
      median_waiting = round(median(waiting_time_days), 1)
    )
  print("Waiting time data available:")
  print(waiting_summary)
} else {
  print("Warning: waiting_time_days column not found. Please regenerate data using updated match_run_data_prep.Rmd")
}

print("Ready for National US-CRS allocation")
```

# National US-CRS Allocation Functions
```{r allocation-functions}
# Function to determine blood type compatibility (handles A1, A2, A1B, A2B variants)
is_blood_compatible <- function(donor_abo, candidate_abo) {
  
  # Normalize blood types to basic ABO groups
  normalize_abo <- function(abo) {
    case_when(
      abo %in% c("A", "A1", "A2") ~ "A",
      abo %in% c("B") ~ "B", 
      abo %in% c("AB", "A1B", "A2B") ~ "AB",
      abo %in% c("O") ~ "O",
      TRUE ~ abo  # Keep original if not recognized
    )
  }
  
  # Normalize both donor and candidate ABO
  donor_norm <- normalize_abo(donor_abo)
  candidate_norm <- normalize_abo(candidate_abo)
  
  # Apply compatibility rules
  case_when(
    donor_norm == "O" & candidate_norm %in% c("O", "B") ~ "primary",
    donor_norm == "O" & candidate_norm %in% c("A", "AB") ~ "secondary", 
    donor_norm == "A" & candidate_norm %in% c("A", "AB") ~ "primary",
    donor_norm == "B" & candidate_norm %in% c("B", "AB") ~ "primary",
    donor_norm == "AB" & candidate_norm == "AB" ~ "primary",
    TRUE ~ "incompatible"
  )
}

# Function to assign national US-CRS classifications (24 main classes)
assign_national_uscrs_classification <- function(match_data) {
  
  match_data %>%
    mutate(
      # Determine donor age category
      donor_age_category = ifelse(DON_AGE >= 18, "adult", "pediatric"),
      
      # Determine candidate age category and pediatric status mapping
      candidate_age_category = case_when(
        str_detect(status, "pediatric") ~ "pediatric",
        PTR_STAT_CD %in% c("2010", "2020", "2030") ~ "pediatric",
        CAN_AGE_AT_LISTING < 18 ~ "pediatric",
        TRUE ~ "adult"
      ),
      
      # Blood type compatibility
      blood_compatibility = is_blood_compatible(DON_ABO, CAN_ABO),
      
      # US-CRS status categories for adults
      uscrs_status_numeric = case_when(
        us_crs_score >= 49 ~ 1,
        us_crs_score >= 40 & us_crs_score < 49 ~ 2,
        us_crs_score >= 30 & us_crs_score < 40 ~ 3,
        us_crs_score >= 20 & us_crs_score < 30 ~ 4,
        us_crs_score >= 10 & us_crs_score < 20 ~ 5,
        us_crs_score >= 0 & us_crs_score < 10 ~ 6,
        TRUE ~ NA_real_
      ),
      
      # Pediatric status extraction
      pediatric_status = case_when(
        status == "Status 1A (pediatric)" ~ "1A",
        status == "Status 1B (pediatric)" ~ "1B", 
        status == "Status 2 (pediatric)" ~ "2",
        TRUE ~ NA_character_
      )
    ) %>%
    filter(blood_compatibility != "incompatible") %>%
    mutate(
      # National US-CRS allocation classification (24 main classes)
      national_uscrs_classification = case_when(
        
        # TABLE N1: Adult Donors (≥18 years) - Classes 1-12
        donor_age_category == "adult" ~ case_when(
          # Class 1-2: Status 1 + Ped 1A (Highest Priority)
          (candidate_age_category == "adult" & uscrs_status_numeric == 1 |
           candidate_age_category == "pediatric" & pediatric_status == "1A") & 
           blood_compatibility == "primary" ~ 1,
          (candidate_age_category == "adult" & uscrs_status_numeric == 1 |
           candidate_age_category == "pediatric" & pediatric_status == "1A") & 
           blood_compatibility == "secondary" ~ 2,
          
          # Class 3-4: Status 2
          candidate_age_category == "adult" & uscrs_status_numeric == 2 & 
           blood_compatibility == "primary" ~ 3,
          candidate_age_category == "adult" & uscrs_status_numeric == 2 & 
           blood_compatibility == "secondary" ~ 4,
          
          # Class 5-6: Status 3 + Ped 1B
          (candidate_age_category == "adult" & uscrs_status_numeric == 3 |
           candidate_age_category == "pediatric" & pediatric_status == "1B") & 
           blood_compatibility == "primary" ~ 5,
          (candidate_age_category == "adult" & uscrs_status_numeric == 3 |
           candidate_age_category == "pediatric" & pediatric_status == "1B") & 
           blood_compatibility == "secondary" ~ 6,
          
          # Class 7-8: Status 4
          candidate_age_category == "adult" & uscrs_status_numeric == 4 & 
           blood_compatibility == "primary" ~ 7,
          candidate_age_category == "adult" & uscrs_status_numeric == 4 & 
           blood_compatibility == "secondary" ~ 8,
          
          # Class 9-10: Status 5  
          candidate_age_category == "adult" & uscrs_status_numeric == 5 & 
           blood_compatibility == "primary" ~ 9,
          candidate_age_category == "adult" & uscrs_status_numeric == 5 & 
           blood_compatibility == "secondary" ~ 10,
          
          # Class 11-12: Status 6 + Ped 2 (Lowest Priority)
          (candidate_age_category == "adult" & uscrs_status_numeric == 6 |
           candidate_age_category == "pediatric" & pediatric_status == "2") & 
           blood_compatibility == "primary" ~ 11,
          (candidate_age_category == "adult" & uscrs_status_numeric == 6 |
           candidate_age_category == "pediatric" & pediatric_status == "2") & 
           blood_compatibility == "secondary" ~ 12,
          
          TRUE ~ 999
        ),
        
        # TABLE N2: Pediatric Donors (<18 years) - Classes 13-24+
        donor_age_category == "pediatric" ~ case_when(
          # Class 13-14: Pediatric Status 1A (PEDIATRIC PRIORITY)
          candidate_age_category == "pediatric" & pediatric_status == "1A" & 
           blood_compatibility == "primary" ~ 13,
          candidate_age_category == "pediatric" & pediatric_status == "1A" & 
           blood_compatibility == "secondary" ~ 14,
          
          # Class 15-16: Adult Status 1 (After pediatric 1A)
          candidate_age_category == "adult" & uscrs_status_numeric == 1 & 
           blood_compatibility == "primary" ~ 15,
          candidate_age_category == "adult" & uscrs_status_numeric == 1 & 
           blood_compatibility == "secondary" ~ 16,
          
          # Class 17-18: Pediatric Status 1B (PEDIATRIC PRIORITY)
          candidate_age_category == "pediatric" & pediatric_status == "1B" & 
           blood_compatibility == "primary" ~ 17,
          candidate_age_category == "pediatric" & pediatric_status == "1B" & 
           blood_compatibility == "secondary" ~ 18,
          
          # Class 19-20: Adult Status 2 (After pediatric 1B)
          candidate_age_category == "adult" & uscrs_status_numeric == 2 & 
           blood_compatibility == "primary" ~ 19,
          candidate_age_category == "adult" & uscrs_status_numeric == 2 & 
           blood_compatibility == "secondary" ~ 20,
          
          # Class 21-22: Adult Status 3
          candidate_age_category == "adult" & uscrs_status_numeric == 3 & 
           blood_compatibility == "primary" ~ 21,
          candidate_age_category == "adult" & uscrs_status_numeric == 3 & 
           blood_compatibility == "secondary" ~ 22,
          
          # Class 23-24: Pediatric Status 2 (PEDIATRIC PRIORITY)  
          candidate_age_category == "pediatric" & pediatric_status == "2" & 
           blood_compatibility == "primary" ~ 23,
          candidate_age_category == "pediatric" & pediatric_status == "2" & 
           blood_compatibility == "secondary" ~ 24,
          
          # Continue with Adult Status 4, 5, 6...
          # For simplicity, remaining classes continue the pattern
          candidate_age_category == "adult" & uscrs_status_numeric %in% c(4, 5, 6) ~ 25 + (uscrs_status_numeric - 4) * 2 + ifelse(blood_compatibility == "secondary", 1, 0),
          
          TRUE ~ 999
        ),
        
        TRUE ~ 999
      )
    )
}

# Function to calculate national US-CRS sequence within each match run
calculate_national_uscrs_sequence <- function(classified_data) {
  
  classified_data %>%
    group_by(MATCH_ID) %>%
    arrange(
      national_uscrs_classification,           # Primary: Classification (lower = higher priority)
      desc(us_crs_score),                     # Secondary: Higher US-CRS score within class
      desc(1 - prob_surv_6wk),                # Tertiary: Lower survival probability (higher risk)
      desc(waiting_time_days)                 # Quaternary: Longer waiting time
    ) %>%
    mutate(
      national_uscrs_sequence = row_number()  # New sequence based on national US-CRS allocation
    ) %>%
    ungroup()
}
```

# Test National Allocation
```{r test-national-allocation}
# Test the national classification and sequencing on a sample match run
sample_match_id <- unique(heart_only_match_run_with_uscrs$MATCH_ID)[1]

# Apply national allocation to sample match run
test_national_data <- heart_only_match_run_with_uscrs %>%
  filter(MATCH_ID == sample_match_id) %>%
  assign_national_uscrs_classification() %>%
  calculate_national_uscrs_sequence()

# Display comparison of original vs national US-CRS sequence
test_national_comparison <- test_national_data %>%
  select(PTR_SEQUENCE_NUM, national_uscrs_sequence, national_uscrs_classification, 
         uscrs_status_numeric, us_crs_score, prob_surv_6wk, 
         waiting_time_days, blood_compatibility, donor_age_category, candidate_age_category) %>%
  arrange(national_uscrs_sequence)

print("Sample National US-CRS Allocation:")
print(test_national_comparison)

# Summary of classification distribution  
cat("\nClassification Distribution:\n")
print(table(test_national_data$national_uscrs_classification))

cat("\nDonor Type:", unique(test_national_data$donor_age_category), "\n")
cat("Blood Types Present:", unique(test_national_data$blood_compatibility), "\n")
```

# Apply National Allocation to Full Dataset  
```{r apply-national-allocation}
print("Starting National US-CRS allocation for full dataset...")

print(paste("Processing", nrow(heart_only_match_run_with_uscrs), "records..."))

# Apply national classification and sequencing
heart_only_match_run_national <- heart_only_match_run_with_uscrs %>%
  assign_national_uscrs_classification() %>%
  calculate_national_uscrs_sequence()

print("National US-CRS allocation completed!")

# Summary statistics
cat("\nNATIONAL ALLOCATION SUMMARY:\n")
cat("- Total records processed:", nrow(heart_only_match_run_national), "\n")
cat("- Records with national classification:", 
    sum(!is.na(heart_only_match_run_national$national_uscrs_classification) & 
        heart_only_match_run_national$national_uscrs_classification != 999), "\n")

# Classification distribution
cat("\nClassification Distribution:\n")
national_class_summary <- heart_only_match_run_national %>%
  filter(national_uscrs_classification != 999) %>%
  count(national_uscrs_classification) %>%
  arrange(national_uscrs_classification)

print(head(national_class_summary, 15))

# Sequence changes compared to original
national_sequence_comparison <- heart_only_match_run_national %>%
  mutate(
    sequence_change = national_uscrs_sequence - PTR_SEQUENCE_NUM,
    sequence_rank_change = case_when(
      sequence_change < 0 ~ "moved_up",
      sequence_change > 0 ~ "moved_down", 
      sequence_change == 0 ~ "same_position",
      TRUE ~ "unknown"
    )
  )

cat("\nSEQUENCE CHANGES (National vs Original):\n")
national_sequence_summary <- national_sequence_comparison %>%
  count(sequence_rank_change) %>%
  mutate(percentage = round(n / sum(n) * 100, 1))

print(national_sequence_summary)
```



# Detailed Sequence Analysis
```{r sequence-impact-analysis}
# Analyze the impact of switching to national US-CRS
sequence_impact <- heart_only_match_run_national %>%
  mutate(
    sequence_change = national_uscrs_sequence - PTR_SEQUENCE_NUM,
    improvement_category = case_when(
      sequence_change <= -10 ~ "Major Improvement (≥10 spots up)",
      sequence_change <= -5 ~ "Moderate Improvement (5-9 spots up)", 
      sequence_change <= -1 ~ "Minor Improvement (1-4 spots up)",
      sequence_change == 0 ~ "No Change",
      sequence_change <= 4 ~ "Minor Decline (1-4 spots down)",
      sequence_change <= 9 ~ "Moderate Decline (5-9 spots down)",
      sequence_change >= 10 ~ "Major Decline (≥10 spots down)"
    ),
    improvement_category = factor(improvement_category, levels = c(
      "Major Improvement (≥10 spots up)",
      "Moderate Improvement (5-9 spots up)", 
      "Minor Improvement (1-4 spots up)",
      "No Change",
      "Minor Decline (1-4 spots down)",
      "Moderate Decline (5-9 spots down)",
      "Major Decline (≥10 spots down)"
    ))
  )

# Impact by US-CRS score
p2 <- sequence_impact %>%
  filter(!is.na(us_crs_score) & !is.na(improvement_category)) %>%
  ggplot(aes(x = us_crs_score, fill = improvement_category)) +
  geom_histogram(binwidth = 2, alpha = 0.8, color = "white", size = 0.1) +
  scale_fill_brewer(type = "div", palette = "RdYlGn", name = "Sequence Change", 
                    guide = guide_legend(reverse = TRUE)) +
  labs(
    title = "Impact of National US-CRS Allocation by Risk Score",
    subtitle = "How sequence position changes vary across US-CRS risk levels",
    x = "US-CRS Score (Higher = Higher Transplant Need)",
    y = "Number of Candidates"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 11, color = "gray40"),
    legend.position = "bottom",
    legend.title = element_text(size = 10, face = "bold")
  ) +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE))

print(p2)

# Save impact analysis
ggsave(here("results", "national_uscrs_impact_by_score.png"), p2, 
       width = 12, height = 8, dpi = 300)

# Summary statistics
impact_summary <- sequence_impact %>%
  filter(!is.na(improvement_category)) %>%
  count(improvement_category) %>%
  mutate(percentage = round(n / sum(n) * 100, 1))

print("Sequence Change Impact Summary:")
print(impact_summary)
```

# Status-Specific Analysis
```{r status-comparison}
# Compare how different status groups are affected
status_impact <- sequence_impact %>%
  filter(!is.na(status) & !is.na(us_crs_score)) %>%
  group_by(status) %>%
  summarise(
    n_candidates = n(),
    mean_uscrs = round(mean(us_crs_score, na.rm = TRUE), 1),
    median_uscrs = round(median(us_crs_score, na.rm = TRUE), 1),
    mean_sequence_change = round(mean(sequence_change, na.rm = TRUE), 1),
    median_sequence_change = round(median(sequence_change, na.rm = TRUE), 1),
    pct_improved = round(sum(sequence_change < 0, na.rm = TRUE) / n() * 100, 1),
    .groups = "drop"
  ) %>%
  arrange(mean_uscrs)

print("Impact by Current Status:")
print(status_impact)

# Visualization by status
p3 <- sequence_impact %>%
  filter(!is.na(status) & !is.na(sequence_change)) %>%
  ggplot(aes(x = status, y = sequence_change, fill = status)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red", alpha = 0.7) +
  scale_fill_viridis_d(option = "turbo", name = "Current Status") +
  labs(
    title = "Sequence Position Changes by Current Heart Status",
    subtitle = "Negative values = improved priority, Positive values = lower priority",
    x = "Current Heart Status",
    y = "Sequence Position Change\n(National US-CRS - Current)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 11, color = "gray40"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )

print(p3)

ggsave(here("results", "national_uscrs_status_impact.png"), p3, 
       width = 10, height = 7, dpi = 300)

cat("Enhanced visualizations saved to results/ directory\n")
```
# Average US-CRS by Sequence Position
```{r}
# Analyze average US-CRS score by sequence position for both systems
heart_only_match_run_national %>%
  select(PX_ID, MATCH_ID, PTR_SEQUENCE_NUM, us_crs_score, national_uscrs_sequence) %>%
  rename(current_sequence = PTR_SEQUENCE_NUM) %>%
  pivot_longer(cols = c(current_sequence, national_uscrs_sequence), 
               names_to = "allocation_system", 
               values_to = "sequence_position") %>%
  mutate(
    allocation_system = case_when(
      allocation_system == "current_sequence" ~ "Current 6-Status System",
      allocation_system == "national_uscrs_sequence" ~ "National US-CRS System"
    ),
    allocation_system = factor(allocation_system, levels = c("Current 6-Status System", "National US-CRS System"))
  ) %>%
  filter(sequence_position <= 200) %>%  # Focus on top 100 candidates for clarity
  group_by(allocation_system, sequence_position) %>%
  summarise(
    mean_us_crs = mean(us_crs_score, na.rm = TRUE),
    median_us_crs = median(us_crs_score, na.rm = TRUE),
    n_candidates = n(),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = sequence_position, y = mean_us_crs, color = allocation_system)) +
  geom_line(size = 1) +
  geom_point(alpha = 0.7) +
  labs(
    title = "Average US-CRS Score by Sequence Position",
    subtitle = "Comparison between Current 6-Status and National US-CRS Systems",
    x = "Sequence Position (Lower = Higher Priority)",
    y = "Average US-CRS Score"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 11, color = "gray40"),
    legend.position = "right"
  )
```


# Save National Allocation Results
```{r save-national-results}
# Save the complete national allocation dataset
save(heart_only_match_run_national, 
     file = here("data", "heart_only_match_run_national.Rdata"))

# Create summary dataset with key variables
national_summary <- heart_only_match_run_national %>%
  select(
    # Original identifiers
    MATCH_ID, DONOR_ID, PX_ID, DON_AGE,
    
    # Original sequence and new national US-CRS sequence  
    PTR_SEQUENCE_NUM, national_uscrs_sequence,
    
    # Classification variables
    national_uscrs_classification, uscrs_status_numeric, 
    donor_age_category, candidate_age_category,
    
    # Scores and tie-breakers
    us_crs_score, prob_surv_6wk, waiting_time_days,
    
    # Match characteristics
    blood_compatibility, PTR_DISTANCE,
    
    # Original status
    status, PTR_STAT_CD,
    
    # Outcome
    REC_TX_DT
  ) %>%
  mutate(
    # Calculate changes
    sequence_change = national_uscrs_sequence - PTR_SEQUENCE_NUM,
    received_transplant = !is.na(REC_TX_DT)
  )

# Save summary
write_csv(national_summary, 
          here("data", "national_allocation_summary.csv"))

print("National allocation results saved!")
print("Files created:")
print("1. heart_only_match_run_national.Rdata - Complete national dataset")
print("2. national_allocation_summary.csv - Summary with key variables")
print("")
print("Key Features:")
print("- No distance constraints (national allocation)")
print("- 24+ simplified classifications")
print("- Pediatric priority for pediatric donors maintained")
print("- Same US-CRS and tie-breaking logic as full system")
```