---
title: "SRTR Figures Template"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
# Load essential libraries
library(tidyverse)
library(haven)  # For reading SRTR SAS files
library(lubridate)  # For date handling

```
# Table 1 Example
```{r}
# select relevant variables, stratify by one, relabel variable names and specify statistics types
heart_adi_table = heart_SVM |> 
  select(CAN_AGE_AT_LISTING, Male, can_race, adi_decile, durable_LVAD) |> 
  filter(adi_decile == 1 | adi_decile == 10) |> 
  group_by(PX_ID) |> 
  tbl_summary(by = adi_decile, label = list(
    can_race ~ "Race",
    CAN_AGE_AT_LISTING ~ "Age",
    durable_LVAD ~ "Durable LVAD",    
  ),
  statistic = list(all_continuous() ~ "{mean} ({sd})", all_categorical() ~ "{n} ({p}%)"))

```

# Forest Plot Example (for a competing risks regression output)
```{r}
crr_model =  crr(Surv(survival_time,status) ~ cov1, data = data)
forest <- tidy(crr_model)
forest <- forest |> 
  mutate(hazard = exp(estimate),
         conf_int=1.96 * std.error,
         conf_low = exp(estimate - conf_int),
         conf_high = exp(estimate + conf_int))


forest_plot <- forest |> 
   ggplot(aes(x = hazard, y = cov1)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf_low,xmax=conf_high))+
  geom_vline(xintercept = 1, linetype = "dashed", color = "black")+
  xlim(0.5,1.2)+
  labs(x = "",
       y = "cov1",
       title = "Unadjusted Model")+
  theme_classic()
```

# Strobe diagram example (2 exclusions, 5 outcomes)
```{r}
library(DiagrammeR)
library(glue)


strobe = grViz(
  glue("digraph my_flowchart {{ 
      graph[splines = ortho]
      node [fontname = Helvetica, shape = box, width = 4, height = 1]
      
        // Main node
        node1[label = <Kidney Transplant Candidate Registrations, 2014-2022 (n = 362,752)>]
                
        // Invisible blank node to adjust positioning
        blank1[label = '', width = 0.01, height = 0.01]
        
        // Excluded criteria (shift to left)
        excluded1[label = <Pediatric Candidates (n = 9,513)>]
        
        // Connect Total available patients to blank1 (no arrow head)
        node1 -> blank1[dir = none];
        
        // Connect blank1 to excluded1 (on the left side)
        excluded1 -> blank1[dir = back, minlen = 2];
        
        // Keep excluded1 on the left of blank1
        {{ rank = same; blank1 excluded1 }}

        // Blank node for the next exclusion step
        blank2[label = '', width = 0.01, height = 0.01]
        
        // Excluded due to missing values (same rank)
        excluded2[label = <Multiple Listings (n = 40,272)>]
        
        // Connect blank1 to blank2 (no arrow head)
        blank1 -> blank2[dir = none];
        
        // Connect blank2 to excluded2 (right side)
        blank2 -> excluded2[minlen = 2];
        
        // Keep blank2 and excluded2 aligned in the same rank
        {{ rank = same; blank2 excluded2 }}
        
        // Connect blank2 to next step (included for analysis)
        node2[label = <Adult Kidney Transplant Candidate Registrations (n = 312,967)>]
        blank2 -> node2;
        
        // Connect to data linking
        node3[label = <Received Kidney Transplant (DDKT) (n = 114,103)>]
        node4[label = <Died (n = 25,504)>]
        node5[label = <Censored (n = 68,533)>]
        node6[label = <Living Donor kidney transplant (n = 48,289)>]
        node7[label = <Removed for Other Reason (n = 56,538)>]
        
        node2 -> node3;
        node2 -> node4;
        node2 -> node5;
        node2 -> node6;
        node2 -> node7;
     }}")
)
```

# CIF plot example with two variables to stratify by, multiple competing risks
```{r}
cuminc_obj <- tidycmprsk::cuminc(Surv(time = time, event = as.factor(status)) ~ cov1 + cov2, 
                                 data = df)

data = cuminc_obj$tidy %>% filter(outcome == 'death')
data = data %>%
  separate(strata, into = c("cov1", "cov2"), sep = ", ") %>% mutate(
    cov1 = as.factor(as.numeric(cov1)),
    cov2 =  as.factor(as.numeric(cov2))
  )

#plot with faceting by cov2

ggplot(data, aes(x = time, y = estimate, color = cov1)) +
  geom_line() +
  facet_wrap(~ cov2)  +
  scale_color_manual(values = blue_palette) + 
  theme_bw()
```

# Kaplan Meier Curve Example
```{r}
fit = survfit( Surv(time, event)~ as.factor(type), data=wgt_diff_8)
cut_down_label = function(x){
  substr(x, 6, nchar(x))
}
summary_fit = tibble(counts = fitfit$strata,
                                 type = cut_down_label(names(fit$strata))
) %>% 
  uncount(counts) %>% 
  cbind(
    tibble(time = fit$time,
           survival = fit$surv,
           lower = fit$lower,
           upper = fit$upper)
  ) %>%
  filter(time <= 1825)  # five year survival

# For a plot stratified by 6 groups
# Includes examples of how to customize axes / colors
colors = scales::hue_pal()(6)
fill_labels = c("Group 1", Group 2", "Group 3", "Group 4", "Group 5", "Group 6")
summary_fit %>%
  ggplot(aes(x= time/365, y = survival, ymin = lower, ymax = upper,color = type)) +
  geom_step() + 
  scale_x_continuous(breaks = seq(0, 5, 1)) +
  labs(color = "", y = "Graft Survival", x = "Time (years post transplant)") +  # Moved labs() inside ggplot()
  ggthemes::theme_gdocs() + lims(y = c(0.75, 1), x = c(0, 1.005)) +
  theme(
    plot.title = element_text(color = "black"),  # Title color
    axis.title.x = element_text(color = "black"),  # X-axis label color
    axis.title.y = element_text(color = "black"),  # Y-axis label color
    axis.text.x = element_text(color = "black"),  # X-axis tick labels color
    axis.text.y = element_text(color = "black"),  # Y-axis tick labels color
    legend.text = element_text(color = "black"),  # Legend labels color
    legend.title = element_text(color = "black"))+   # Legend title color
scale_color_manual(values = colors, labels = fill_labels) 

